\section{Problem Encoding}

A molecule consists of a number of atoms. Each of these atoms has its own 3-dimensional position within the molecule. For the structure refinement problem the individual 3-dimensional position values are not important. The important information about this problem is how the atoms are positioned with respect to each other. Two different forms of representation were used in this work. For each of these representations the number values are shown in Angstroms (\AA).

The initial run of experiments used a representation that maintained the initial atomic positions of each atom. The 3-dimensional coordinates were treated as a list of coordinates as shown in~\ref{fig:representation1}. Using this representation meant that during any form of crossover the tuple of X, Y and Z values would stay together.

\begin{figure}
	\centering
	\begin{tabular}{ | l | l | l | }
		\hline
		X & Y & Z \\ \hline
		14.451 & -13.346 & 1.133 \\ \hline
		15.336 & -13.488 & 2.014 \\ \hline
		13.005 & -13.364 & 1.452 \\ \hline
		0.019 & 0.011 & 0.045 \\ \hline
		... & ... & ... \\ \hline
	\end{tabular}
	\caption{Representation 1}
	\label{fig:representation1}
\end{figure}

The other representation used was a list of values. The initial list of 3-dimentional coordinates was converted to a single list of decimal points as shown in~\ref{fig:representation2}. Algorithms such as particle swarm optimization and differential evolution called for a more flexible representation. For fitness evaluation the list of number was converted back to a list of 3-dimentional coordinates by taking segments of three numbers to create a 3-dimentional position.

\begin{figure}
	\centering
	\begin{tabular}{ | l | }
		\hline
		14.451 \\ \hline
		-13.346 \\ \hline
		1.133 \\ \hline
		15.336 \\ \hline
		-13.488 \\ \hline
		2.014 \\ \hline
		13.005 \\ \hline
		... \\ \hline
	\end{tabular}
	\caption{Representation 2}
	\label{fig:representation2}
\end{figure}